Bayesian Statistical Modeling with Predictors from LLMs

### 2024-04-05 04-paper/paper-LLM-probmod.tex
revised secs 5-6, C, tables and figs

\paragraph{Notation.}
Let $\set{I_{1}, \dots, I_{m}}$ be $m$ be instances of the same task, or items belonging to the same (logical) condition in a behavioral experiment.
% For human observers, these items are interchangeable, all exemplifying the same conceptual problem.
Each item $I_{k} = \tuple{x_{k}, \tuple{y_{k1}, \dots, y_{kl}}}$ consists of an input prompt $x_{k}$, which is a string of text, and $l$ choice options $\tuple{y_{k1}, \dots, y_{kl}}$, all of which are strings as well, possibly composed of $|y_{ki}|$ tokens, $y_{ki} = w_{ki1}, \dots, w_{ki|y_{ki}|}$.
For simplicity of notation, we assume that the $i$-th choice option $y_{ki}$ for each item $k$ belongs to the same response category $R_{i}$.
For the case at hand these categories are: target, competitor, distractor, so that $y_{k1}$ always corresponds to the designated \emph{target option}, $R_{1}$.

\paragraph{LLM scores \& predictions.}
The most obvious \emph{item-level score} an (autoregressive) LLM provides for each choice option $y_{ki}$ is its log-probability:\footnote{
  More elaborate item-level scores include corrections for variable length of answer options \citep[e.g.,][]{BrownMann2020:Language-Models} or variation in base rate among answer options \citep[e.g.,][]{HoltzmanWest2021:Surface-Form-Co}.
  From the point of view of experimental psychology, these corrections are \emph{post hoc} fixes to improperly balanced experimental materials.
  For the purposes of this paper, where answer options are all equally long and commensurable, these corrections may be temporarily ignored for simplicity.
  \mf{@PT: check whether this is correct also after new LLaMA results.}
}
%
\begin{align*}
  % S\left( y_{ki}, x_{k} \right) =
  \text{S}_{ki}
  % = \log P_{\text{LLM}} \left(y_{ki} \mid x_{k} \right)
  =  \sum_{j=1}^{|y_{ki}|} \log P_{\text{LLM}} \left(w_{kij} \mid x_{k}, w_{ki1}, \dots, w_{ki(j-1)} \right)  \,.
\end{align*}
%

Based on each option's score, we can define the \emph{item-level prediction} of choosing option $y_{ki}$ in term of soft-maximization as:
%
\begin{align*}
P_{\text{item}}^{\text{LLM}}\left ( y_{ki} \mid C; \alpha_{c} \right ) \propto \expo \left (\alpha_{c} \ \text{S}_{ki} \right)\,.
\end{align*}
%
Notice that the usual item-level prediction used in benchmark testing is actually a ``winner-takes-all'' strategy, but this is just a special case of the above for $\alpha_{C} \rightarrow \infty$.
Here, the $\alpha_{C}$ parameter corresponds to inverse temperature, so that $\alpha_{C} \rightarrow \infty$ corresponds to greedy sampling with temperature zero.
Like in the RSA model, we use an independent soft-max parameter for each condition.

@@

To address this question, we assessed item-level scores $\text{S}_{ki}$ from the text-davinci-003 instance GPT-3.5 August 2023 \citep{BrownMann2020:Language-Models} for a text-based version of the reference game from Section~\ref{experiment-reference-games}.
% \footnote{
  % Numerically, we worked with \emph{average} log-probabilities. As in our experiment, all choice options had the same number of words, this is largely irrelevant, except for the effect of the $\alpha$ parameter in the interpretation condition. Using length-corrected log-probabilities is the same as using raw log-probabilities with an $\alpha$ parameter divided by the number of words. In this way, we make the $\alpha$ parameter more meaningfully comparable across conditions.}
The task description $x_{k}$ for item $I_{k}$ is a text supplied as prompt (see Appendix~\ref{sec:examples-items-llm} for an example).
The choice options are categorized as in the human experiment.\footnote{
  In the current set-up the response type ``distractor'' has two instantiations in the production condition. Since choice options are a single word in the production condition, for simplicity, we lump both of the distractor words together and treat them as a single option.}
Similar to the RSA model fit, we allow for random errors with probability $\epsilon_{C}$:
%
\begin{align*}
  P_{\text{item}}^{\text{LLM}}\left ( y_{ki} \mid C; \alpha_{c}, \epsilon_{c} \right )
  = \epsilon_{c} \  P_{\text{item}}^{\text{LLM}}\left ( y_{ki} \mid C; \alpha_{c}, \epsilon_{c} \right ) \ \ + \ \ \nicefrac{\epsilon_{c}}{3}   \,.
\end{align*}
%
The resulting likelihood function for item-level data with item-level predictions is, where priors on parameters are as for the RSA model described in Section~\ref{sec:model-pred-from}:
%
\begin{align*}
 P^{\text{LLM}}_{\text{item}}(D_{c} \mid C; \alpha_{c}, \epsilon_{c}) = \prod_{k = 1}^{m} \text{Multinomial} \left ( D_{c}^{k}, \tuple{P^{\text{LLM}}_{\text{item}}(y_{ki} \mid C;  \alpha_{c}, \epsilon_{c})}_{1 \le i \le 3} \right )\,.
\end{align*}
%

@@

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{00-pics/measures-overview.png}
  \caption{
    Schematic overview over three different approaches of obtaining condition-level predictions by aggregating item-level information.
    Approaches differ in the kind of item-level information they take into account, which entails differences in the order in which aggregation, transformation to probabilities and parameterized scaling occur.
  }
  \label{fig:measures-overview}
\end{figure}

While LLMs do not make conditional-level predictions as such, they can be derived from item-level scores $\text{S}_{ki}$ by averaging over all items belonging to the relevant condition.
There are many ways of averaging item-level information.
Figure~\ref{fig:measures-overview} shows three salient approaches, which differ in what the underlying item-level measure is: the raw scores $\text{S}_{ki}$, the item-level probability derived from it (as used in Section~\ref{sec:item-level-pred}), or the predictions from the winner-takes-all (WTA) strategy commonly used in benchmark testing.

@@

The \emph{average-scores predictor} first aggregates the item-level scores, and then transposes the averages into (scaled) probabilities using the usual parameterized softmax function:\footnote{
  The reported results average over the multi-set \(\set{I_1, \dots I_m}\) of items that occurred in the human experiment for condition $C$.
  By using a multi-set, which may contain a single item multiple times, we produce aggregate predictions for exactly the set of items that the participant group saw, which provides the most fitting counterpart to the human data.}
%
\begin{align*}
  & P_{\text{cond}}^{\text{SCR}}(r_{i} \mid C, \alpha_{c})
    \propto \expo \left [  \alpha_{c} \ \frac{1}{m} \ \sum_{k=1}^{m} \text{S}_{ki}  \right ] \,.
    \tag*{\textcolor{gray}{[average scores (narrow-scope aggregation)]}}
\end{align*}
%
The \emph{average-probabilities predictor} first transposes scores into probabilities with a parameterized softmax function, and only aggregates over items last:
\begin{align*}
  % & \textbf{wide-scope aggregation} \\
  & P_{\text{cond}}^{\text{PRB}}(r_{i} \mid C, \alpha_{c})
    = \frac{1}{m} \ \sum_{k=1}^{m} P_{\text{item}}^{\text{LLM}}(y_{ki} \mid C, \alpha_{c}) \,.
    \tag*{\textcolor{gray}{[average probabilities (wide-scope aggregation)]}}
\end{align*}
%
Finally, the \emph{average-WTA predictor} considers the prediction of the WTA strategy (a softmax with $\alpha \rightarrow \infty$) as the basic item-level unit to aggregate over.
To add parameterized scaling to this method, a power-law transformation is a natural choice:
\begin{align*}
  % & \textbf{intermediate-scope aggregation} \\
  & P_{\text{cond}}^{\text{WTA}}(r_{i} \mid C, \alpha_{c})
    \propto  \left [ \frac{1}{m} \ \sum_{k=1}^{m} P_{\text{item}}^{\text{WTA}}(y_{ki} \mid C) \right ]^{\alpha_{c}}\,,
    \tag*{\textcolor{gray}{[average WTA (intermediate-scope aggregation)]}}
\end{align*}
where $P_{\text{item}}^{\text{WTA}}(y_{ki} \mid C) = \lim_{\alpha \rightarrow \infty} P_{\text{item}}^{\text{LLM}}(y_{ki} \mid C, \alpha)$.
\bigskip

@@

All previous results where obtained for LMM predictions derived from GPT-3.5 turbo.
To investigate whether key results replicate for other LLMs and scales, we ran the same analyses also for scores derived from variants of LLaMA2 \citep{TouvronLavril2023:LLaMA:-Open-and}, in particular the 7, 13 and 70 billion parameter versions of the models fine-tuned with human feedback only (marked as `hf' in the following) and based on human feedback and chat data (marked as `chat').
Table~\ref{tab:sumStats-LLaMA2} shows the relevant summary statistics and Bayesian posterior predictive $p$-values.
Figure~\ref{fig:LLaMA-results-vPPC} shows visual posterior predictive checks for the condition-level data.
The plots show the \emph{differences} between observed counts and the models' posterior predictions for expected counts for each condition and response option.
We find that the average-WTA predictor is consistently able to capture the data from the production condition, except for one of the six models (LLaMA2-chat 70B).
However, the average-WTA predictor recovers the interpretation data for only one of the six models (LLaMA2-chat 13B, see also Table~\ref{tab:sumStats-LLaMA2}).
Interestingly, for LLaMA2-chat 13B also the average-probability predictor is able to recover the interpretation data, unlike for the GPT-3.5.
As for GPT-3.5, no model captured the item-level variance observed in the human data.
In sum, we find general support for the previous conclusions that the average-WTA predictor is most successful in capturing the condition-level data, and therefore that, where models manage to recover the condition-level data, they seem to be ``right for the wrong reasons.''
Additionally, we also find variability in which method of aggregation works well with which LLM as scoring model.

\begin{table}[t]
\centering
\begin{tabular}{cllllcrlcrlc}
  \toprule
  &&&& $\alpha$ &&& $\epsilon$ & \\ \cmidrule(r){4-6} \cmidrule(l){7-9}
  & model & data & method & condition & |95\% & mean\ & 95\%| & |95\% & mean\ & 95\%| & Bpppv \\
  \midrule

    & L2-hf-7b    & item  & ---         & prd. & 1.09  & 1.30   & 1.51   & 0.00 & 0.02 & 0.06 & 0.00 \\
    & L2-hf-7b    & item  & ---         & int. & 1.90  & 2.52   & 3.13   & 0.00 & 0.04 & 0.12 & 0.00 \\
  ! & L2-hf-7b    & cond. & avg. scores & prd. & 9.65  & 11.50  & 13.62  & 0.08 & 0.12 & 0.16 & 0.48 \\
    & L2-hf-7b    & cond. & avg. scores & int. & 6.80  & 7.96   & 9.14   & 0.00 & 0.02 & 0.05 & 0.00 \\
    & L2-hf-7b    & cond. & avg. prob.  & prd. & 4.98  & 16.44  & 33.06  & 0.06 & 0.10 & 0.13 & 0.00 \\
    & L2-hf-7b    & cond. & avg. prob.  & int. & 20.06 & 79.30  & 163.94 & 0.00 & 0.01 & 0.04 & 0.00 \\
  ! & L2-hf-7b    & cond. & avg. WTA    & prd. & 4.17  & 4.96   & 5.85   & 0.07 & 0.12 & 0.16 & 0.48 \\
    & L2-hf-7b    & cond. & avg. WTA    & int. & 0.87  & 1.02   & 1.18   & 0.00 & 0.02 & 0.05 & 0.00 \\ \addlinespace[0.5em]
    & L2-hf-13b   & item  & ---         & prd. & 1.37  & 1.91   & 2.47   & 0.00 & 0.07 & 0.12 & 0.00 \\
    & L2-hf-13b   & item  & ---         & int. & 2.13  & 2.90   & 3.69   & 0.00 & 0.05 & 0.14 & 0.00 \\
  ! & L2-hf-13b   & cond. & avg. scores & prd. & 13.73 & 16.32  & 19.10  & 0.08 & 0.12 & 0.16 & 0.50 \\
    & L2-hf-13b   & cond. & avg. scores & int. & 7.73  & 9.11   & 10.42  & 0.00 & 0.02 & 0.05 & 0.00 \\
    & L2-hf-13b   & cond. & avg. prob.  & prd. & 69.84 & 185.80 & 360.44 & 0.06 & 0.09 & 0.13 & 0.00 \\
    & L2-hf-13b   & cond. & avg. prob.  & int. & 13.63 & 39.72  & 88.28  & 0.00 & 0.02 & 0.05 & 0.00 \\
  ! & L2-hf-13b   & cond. & avg. WTA    & prd. & 2.28  & 2.70   & 3.18   & 0.08 & 0.12 & 0.16 & 0.51 \\
    & L2-hf-13b   & cond. & avg. WTA    & int. & 0.78  & 0.94   & 1.12   & 0.00 & 0.03 & 0.07 & 0.03 \\ \addlinespace[0.5em]
    & L2-hf-70b   & item  & ---         & prd. & 1.53  & 1.86   & 2.22   & 0.00 & 0.03 & 0.07 & 0.00 \\
    & L2-hf-70b   & item  & ---         & int. & 3.20  & 4.01   & 4.74   & 0.00 & 0.03 & 0.08 & 0.00 \\
  ! & L2-hf-70b   & cond. & avg. scores & prd. & 44.57 & 53.17  & 62.44  & 0.08 & 0.12 & 0.17 & 0.48 \\
    & L2-hf-70b   & cond. & avg. scores & int. & 6.86  & 8.14   & 9.32   & 0.00 & 0.02 & 0.05 & 0.00 \\
    & L2-hf-70b   & cond. & avg. prob.  & prd. & 10.51 & 26.59  & 45.73  & 0.07 & 0.10 & 0.14 & 0.00 \\
    & L2-hf-70b   & cond. & avg. prob.  & int. & 10.86 & 23.63  & 42.16  & 0.00 & 0.03 & 0.09 & 0.00 \\
  ! & L2-hf-70b   & cond. & avg. WTA    & prd. & 5.88  & 7.03   & 8.19   & 0.08 & 0.12 & 0.17 & 0.50 \\
    & L2-hf-70b   & cond. & avg. WTA    & int. & 0.55  & 0.68   & 0.81   & 0.00 & 0.03 & 0.07 & 0.04 \\ \addlinespace[0.5em]
    & L2-chat-7b  & item  & ---         & prd. & 0.34  & 0.41   & 0.49   & 0.00 & 0.03 & 0.07 & 0.00 \\
    & L2-chat-7b  & item  & ---         & int. & 0.79  & 1.05   & 1.34   & 0.00 & 0.05 & 0.15 & 0.00 \\
  ! & L2-chat-7b  & cond. & avg. scores & prd. & 58.47 & 69.64  & 81.79  & 0.08 & 0.12 & 0.16 & 0.48 \\
    & L2-chat-7b  & cond. & avg. scores & int. & 2.03  & 2.37   & 2.73   & 0.00 & 0.02 & 0.05 & 0.00 \\
    & L2-chat-7b  & cond. & avg. prob.  & prd. & 8.05  & 27.26  & 55.30  & 0.07 & 0.11 & 0.15 & 0.00 \\
    & L2-chat-7b  & cond. & avg. prob.  & int. & 2.86  & 19.24  & 51.41  & 0.00 & 0.01 & 0.04 & 0.00 \\
  ! & L2-chat-7b  & cond. & avg. WTA    & prd. & 12.18 & 14.70  & 17.21  & 0.08 & 0.12 & 0.16 & 0.50 \\
    & L2-chat-7b  & cond. & avg. WTA    & int. & 0.86  & 1.01   & 1.14   & 0.00 & 0.02 & 0.05 & 0.00 \\ \addlinespace[0.5em]
    & L2-chat-13b & item  & ---         & prd. & 0.36  & 0.44   & 0.53   & 0.00 & 0.03 & 0.08 & 0.00 \\
    & L2-chat-13b & item  & ---         & int. & 0.68  & 0.90   & 1.11   & 0.00 & 0.05 & 0.13 & 0.00 \\
  ! & L2-chat-13b & cond. & avg. scores & prd. & 5.56  & 6.63   & 7.76   & 0.08 & 0.12 & 0.16 & 0.49 \\
    & L2-chat-13b & cond. & avg. scores & int. & 1.70  & 2.05   & 2.40   & 0.00 & 0.02 & 0.06 & 0.01 \\
    & L2-chat-13b & cond. & avg. prob.  & prd. & 3.90  & 22.25  & 74.12  & 0.06 & 0.10 & 0.14 & 0.00 \\
  ! & L2-chat-13b & cond. & avg. prob.  & int. & 2.85  & 4.70   & 7.01   & 0.00 & 0.03 & 0.07 & 0.08 \\
  ! & L2-chat-13b & cond. & avg. WTA    & prd. & 4.28  & 5.11   & 5.99   & 0.07 & 0.12 & 0.16 & 0.49 \\
  ! & L2-chat-13b & cond. & avg. WTA    & int. & 0.15  & 0.32   & 0.51   & 0.02 & 0.11 & 0.18 & 0.50 \\ \addlinespace[0.5em]
    & L2-chat-70b & item  & ---         & prd. & 0.28  & 0.35   & 0.42   & 0.00 & 0.03 & 0.08 & 0.00 \\
    & L2-chat-70b & item  & ---         & int. & 1.03  & 1.24   & 1.47   & 0.00 & 0.03 & 0.08 & 0.00 \\
    & L2-chat-70b & cond. & avg. scores & prd. & 0.40  & 0.55   & 0.74   & 0.00 & 0.04 & 0.10 & 0.00 \\
  ! & L2-chat-70b & cond. & avg. scores & int. & 1.91  & 2.37   & 2.89   & 0.00 & 0.03 & 0.08 & 0.16 \\
    & L2-chat-70b & cond. & avg. prob.  & prd. & 0.43  & 0.66   & 0.96   & 0.00 & 0.04 & 0.10 & 0.00 \\
    & L2-chat-70b & cond. & avg. prob.  & int. & 2.82  & 5.21   & 8.43   & 0.00 & 0.01 & 0.04 & 0.02 \\
    & L2-chat-70b & cond. & avg. WTA    & prd. & 0.14  & 0.21   & 0.35   & 0.00 & 0.06 & 0.13 & 0.00 \\

>>>>

\paragraph{Notation.}
Let $\set{I_{1}, \dots, I_{m}}$ be $m$ be instances of the same task, or items belonging to the same (logical) condition in a behavioral experiment.
% For human observers, these items are interchangeable, all exemplifying the same conceptual problem.
Each item $I_{k} = \tuple{x_{k}, \tuple{y_{kl}}_{1 \le l \le n}}$ consists of an input prompt $x_{k}$, which is a string of text, and $n$ choice options $\tuple{y_{kl}}$, all of which are strings as well, possibly composed of $|y_{kl}|$ tokens, $y_{kl} = w_{kl1}, \dots, w_{kl|y_{kl}|}$.
For simplicity of notation, we assume that the $l$-th choice option $y_{kl}$ for each item $k$ belongs to the same response category $R_{l}$.
For the case at hand these categories are: target, competitor, distractor, so that $y_{k1}$ always corresponds to the designated \emph{target option}, $R_{1}$.

\paragraph{LLM scores \& predictions.}
The most obvious \emph{item-level score} an (autoregressive) LLM provides for each choice option $y_{kl}$ is its log-probability:\footnote{
  More elaborate item-level scores include corrections for variable length of answer options \citep[e.g.,][]{BrownMann2020:Language-Models} or variation in base rate among answer options \citep[e.g.,][]{HoltzmanWest2021:Surface-Form-Co}.
  From the point of view of experimental psychology, these corrections are \emph{post hoc} fixes to improperly balanced experimental materials.
  For the purposes of this paper, where answer options are all equally long and commensurable, these corrections may be temporarily ignored for simplicity.
  \mf{@PT: check whether this is correct also after new LLaMA results.}
}
%
\begin{align*}
  % S\left( y_{ki}, x_{k} \right) =
  \text{S}_{kl}
  % = \log P_{\text{LLM}} \left(y_{ki} \mid x_{k} \right)
  =  \sum_{i=1}^{|y_{kl}|} \log P_{\text{LLM}} \left(w_{kli} \mid x_{k}, w_{kl1}, \dots, w_{kl(i-1)} \right)  \,.
\end{align*}
%

Based on each option's score, we can define the \emph{item-level prediction} of choosing option $y_{ki}$ in term of soft-maximization as:
%
\begin{align*}
P_{\text{item}}^{\text{LLM}}\left ( y_{kl} \mid C; \alpha_{c} \right ) \propto \expo \left (\alpha_{c} \ \text{S}_{kl} \right)\,.
\end{align*}
%
Notice that the usual item-level prediction used in benchmark testing is actually a ``winner-takes-all'' strategy, but this is just a special case of the above for $\alpha_{C} \rightarrow \infty$.
Here, the $\alpha_{C}$ parameter corresponds to inverse temperature, so that $\alpha_{C} \rightarrow \infty$ corresponds to greedy sampling with temperature zero.
Like in the RSA model, we use an independent soft-max parameter for each condition.

@@

To address this question, we assessed item-level scores $\text{S}_{kl}$ from the text-davinci-003 instance GPT-3.5 August 2023 \citep{BrownMann2020:Language-Models} for a text-based version of the reference game from Section~\ref{experiment-reference-games}.
% \footnote{
  % Numerically, we worked with \emph{average} log-probabilities. As in our experiment, all choice options had the same number of words, this is largely irrelevant, except for the effect of the $\alpha$ parameter in the interpretation condition. Using length-corrected log-probabilities is the same as using raw log-probabilities with an $\alpha$ parameter divided by the number of words. In this way, we make the $\alpha$ parameter more meaningfully comparable across conditions.}
The task description $x_{k}$ for item $I_{k}$ is a text supplied as prompt (see Appendix~\ref{sec:examples-items-llm} for an example).
The choice options are categorized as in the human experiment.\footnote{
  In the current set-up the response type ``distractor'' has two instantiations in the production condition. Since choice options are a single word in the production condition, for simplicity, we lump both of the distractor words together and treat them as a single option.}
Similar to the RSA model fit, we allow for random errors with probability $\epsilon_{C}$:
%
\begin{align*}
  P_{\text{item}}^{\text{LLM}}\left ( y_{kl} \mid C; \alpha_{c}, \epsilon_{c} \right )
  = \epsilon_{c} \  P_{\text{item}}^{\text{LLM}}\left ( y_{kl} \mid C; \alpha_{c}, \epsilon_{c} \right ) \ \ + \ \ \nicefrac{\epsilon_{c}}{3}   \,.
\end{align*}
%
The resulting likelihood function for item-level data with item-level predictions is, where priors on parameters are as for the RSA model described in Section~\ref{sec:model-pred-from}:
%
\begin{align*}
 P^{\text{LLM}}_{\text{item}}(D_{c} \mid C; \alpha_{c}, \epsilon_{c}) = \prod_{k = 1}^{m} \text{Multinomial} \left ( D_{c}^{k}, \tuple{P^{\text{LLM}}_{\text{item}}(y_{kl} \mid C;  \alpha_{c}, \epsilon_{c})}_{1 \le l \le 3} \right )\,.
\end{align*}
%

@@

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{00-pics/measures-overview.png}
  \caption{
    Schematic overview over three different methods of obtaining condition-level predictions by aggregating item-level information.
    Methods differ in the kind of item-level information they take into account, which entails differences in the order in which aggregation, transformation to probabilities and parameterized scaling occur.
  }
  \label{fig:measures-overview}
\end{figure}

While LLMs do not make conditional-level predictions as such, they can be derived from item-level scores $\text{S}_{kl}$ by averaging over all items belonging to the relevant condition.
There are many ways of averaging item-level information.
Figure~\ref{fig:measures-overview} shows three salient approaches, which differ in what the underlying item-level measure for aggregation is: the raw scores $\text{S}_{kl}$, the item-level probability derived from it (as used in Section~\ref{sec:item-level-pred}), or the predictions from the winner-takes-all (WTA) strategy commonly used in benchmark testing.

@@

The \emph{average-scores predictor} first aggregates the item-level scores, and then transposes the averages into (scaled) probabilities using the usual parameterized softmax function:\footnote{
  The reported results average over the multi-set \(\set{I_1, \dots I_m}\) of items that occurred in the human experiment for condition $C$.
  By using a multi-set, which may contain a single item multiple times, we produce aggregate predictions for exactly the set of items that the participant group saw, which provides the most fitting counterpart to the human data.}
%
\begin{align*}
  & P_{\text{cond}}^{\text{SCR}}(R_{l} \mid C, \alpha_{c})
    \propto \expo \left [  \alpha_{c} \ \frac{1}{m} \ \sum_{k=1}^{m} \text{S}_{kl}  \right ] \,.
    \tag*{\textcolor{gray}{[average scores (narrow-scope aggregation)]}}
\end{align*}
%
The \emph{average-probabilities predictor} first transposes scores into probabilities with a parameterized softmax function, and only aggregates over items last:
\begin{align*}
  % & \textbf{wide-scope aggregation} \\
  & P_{\text{cond}}^{\text{PRB}}(R_{l} \mid C, \alpha_{c})
    = \frac{1}{m} \ \sum_{k=1}^{m} P_{\text{item}}^{\text{LLM}}(y_{kl} \mid C, \alpha_{c}) \,.
    \tag*{\textcolor{gray}{[average probabilities (wide-scope aggregation)]}}
\end{align*}
%
Finally, the \emph{average-WTA predictor} considers the prediction of the WTA strategy (a softmax with $\alpha \rightarrow \infty$) as the basic item-level unit to aggregate over.
To add parameterized scaling to this method, a power-law transformation is a natural choice:
\begin{align*}
  % & \textbf{intermediate-scope aggregation} \\
  & P_{\text{cond}}^{\text{WTA}}(R_{l} \mid C, \alpha_{c})
    \propto  \left [ \frac{1}{m} \ \sum_{k=1}^{m} P_{\text{item}}^{\text{WTA}}(y_{kl} \mid C) \right ]^{\alpha_{c}}\,,
    \tag*{\textcolor{gray}{[average WTA (intermediate-scope aggregation)]}}
\end{align*}
where $P_{\text{item}}^{\text{WTA}}(y_{kl} \mid C) = \lim_{\alpha \rightarrow \infty} P_{\text{item}}^{\text{LLM}}(y_{kl} \mid C, \alpha)$.
\bigskip

@@

All previous results where obtained for LMM predictions derived from GPT-3.5 turbo.
To investigate whether key results replicate for other LLMs and scales, we ran the same analyses also for scores derived from variants of LLaMA2 \citep{TouvronLavril2023:LLaMA:-Open-and}, in particular the 7, 13 and 70 billion parameter versions of the models fine-tuned with human feedback only (marked as `hf' in the following) and based on human feedback and chat data (marked as `chat').
Table~\ref{tab:sumStats-LLaMA2} shows the relevant summary statistics and Bayesian posterior predictive $p$-values.
Figure~\ref{fig:LLaMA-results-vPPC} shows visual posterior predictive checks for the condition-level data.
The plots show the \emph{differences} between observed counts and the models' posterior predictions for expected counts for each condition and response option.
We find that the average-WTA predictor is consistently able to capture the data from the production condition, except for one of the six models (LLaMA2-chat 70B).
When the average-WTA predictor passes the model checking criterion for the production data, the posteriors for the power-law parameter $\alpha$ are credibly different from 1, thus suggesting that, contrary to the results for GPT-3.5 a mere ``WTA average'' is not a good prediction strategy for the production data for the LLaMA2-based models.
Looking at interpretation, the average-WTA predictor recovers the interpretation data for only one of the six models (LLaMA2-chat 13B, see also Table~\ref{tab:sumStats-LLaMA2}).
Interestingly, for LLaMA2-chat 13B also the average-probability predictor is able to recover the interpretation data, unlike for the GPT-3.5.
Finally, for item-level data, as for GPT-3.5, no LLaMA2-based model captured the item-level variance observed in the human data.
In sum, we find general support for the previous conclusions that the average-WTA predictor is most successful in capturing the condition-level data, and therefore that, where models manage to recover the condition-level data, they seem to be ``right for the wrong reasons.''
Additionally, we also find variability in which method of aggregation works well with which LLM as scoring model.

\begin{table}[]
\centering
\begin{tabular}{cllllcrlcrlc}
  \toprule
  &&&&&& $\alpha$ &&& $\epsilon$ & \\ \cmidrule(r){6-8} \cmidrule(l){9-11}
  & model & data & method & condition & |95\% & mean\ & 95\%| & |95\% & mean\ & 95\%| & Bpppv \\
  \midrule

  \textcolor{shimmer-main}{$\times$} & L2-hf-7b    & item  & ---         & prd. & 1.09  & 1.30   & 1.51   & 0.00 & 0.02 & 0.06 & 0.00 \\
  \textcolor{shimmer-main}{$\times$} & L2-hf-7b    & item  & ---         & int. & 1.90  & 2.52   & 3.13   & 0.00 & 0.04 & 0.12 & 0.00 \\
  \textcolor{fern-main}{$\checkmark$} & L2-hf-7b    & cond. & avg. scores & prd. & 9.65  & 11.50  & 13.62  & 0.08 & 0.12 & 0.16 & 0.48 \\
  \textcolor{shimmer-main}{$\times$} & L2-hf-7b    & cond. & avg. scores & int. & 6.80  & 7.96   & 9.14   & 0.00 & 0.02 & 0.05 & 0.00 \\
  \textcolor{shimmer-main}{$\times$} & L2-hf-7b    & cond. & avg. prob.  & prd. & 4.98  & 16.44  & 33.06  & 0.06 & 0.10 & 0.13 & 0.00 \\
  \textcolor{shimmer-main}{$\times$} & L2-hf-7b    & cond. & avg. prob.  & int. & 20.06 & 79.30  & 163.94 & 0.00 & 0.01 & 0.04 & 0.00 \\
  \textcolor{fern-main}{$\checkmark$} & L2-hf-7b    & cond. & avg. WTA    & prd. & 4.17  & 4.96   & 5.85   & 0.07 & 0.12 & 0.16 & 0.48 \\
  \textcolor{shimmer-main}{$\times$} & L2-hf-7b    & cond. & avg. WTA    & int. & 0.87  & 1.02   & 1.18   & 0.00 & 0.02 & 0.05 & 0.00 \\ \addlinespace[0.5em]
  \textcolor{shimmer-main}{$\times$} & L2-hf-13b   & item  & ---         & prd. & 1.37  & 1.91   & 2.47   & 0.00 & 0.07 & 0.12 & 0.00 \\
  \textcolor{shimmer-main}{$\times$} & L2-hf-13b   & item  & ---         & int. & 2.13  & 2.90   & 3.69   & 0.00 & 0.05 & 0.14 & 0.00 \\
  \textcolor{fern-main}{$\checkmark$} & L2-hf-13b   & cond. & avg. scores & prd. & 13.73 & 16.32  & 19.10  & 0.08 & 0.12 & 0.16 & 0.50 \\
  \textcolor{shimmer-main}{$\times$} & L2-hf-13b   & cond. & avg. scores & int. & 7.73  & 9.11   & 10.42  & 0.00 & 0.02 & 0.05 & 0.00 \\
  \textcolor{shimmer-main}{$\times$} & L2-hf-13b   & cond. & avg. prob.  & prd. & 69.84 & 185.80 & 360.44 & 0.06 & 0.09 & 0.13 & 0.00 \\
  \textcolor{shimmer-main}{$\times$} & L2-hf-13b   & cond. & avg. prob.  & int. & 13.63 & 39.72  & 88.28  & 0.00 & 0.02 & 0.05 & 0.00 \\
  \textcolor{fern-main}{$\checkmark$} & L2-hf-13b   & cond. & avg. WTA    & prd. & 2.28  & 2.70   & 3.18   & 0.08 & 0.12 & 0.16 & 0.51 \\
  \textcolor{shimmer-main}{$\times$} & L2-hf-13b   & cond. & avg. WTA    & int. & 0.78  & 0.94   & 1.12   & 0.00 & 0.03 & 0.07 & 0.03 \\ \addlinespace[0.5em]
  \textcolor{shimmer-main}{$\times$} & L2-hf-70b   & item  & ---         & prd. & 1.53  & 1.86   & 2.22   & 0.00 & 0.03 & 0.07 & 0.00 \\
  \textcolor{shimmer-main}{$\times$} & L2-hf-70b   & item  & ---         & int. & 3.20  & 4.01   & 4.74   & 0.00 & 0.03 & 0.08 & 0.00 \\
  \textcolor{fern-main}{$\checkmark$} & L2-hf-70b   & cond. & avg. scores & prd. & 44.57 & 53.17  & 62.44  & 0.08 & 0.12 & 0.17 & 0.48 \\
  \textcolor{shimmer-main}{$\times$} & L2-hf-70b   & cond. & avg. scores & int. & 6.86  & 8.14   & 9.32   & 0.00 & 0.02 & 0.05 & 0.00 \\
  \textcolor{shimmer-main}{$\times$} & L2-hf-70b   & cond. & avg. prob.  & prd. & 10.51 & 26.59  & 45.73  & 0.07 & 0.10 & 0.14 & 0.00 \\
  \textcolor{shimmer-main}{$\times$} & L2-hf-70b   & cond. & avg. prob.  & int. & 10.86 & 23.63  & 42.16  & 0.00 & 0.03 & 0.09 & 0.00 \\
  \textcolor{fern-main}{$\checkmark$} & L2-hf-70b   & cond. & avg. WTA    & prd. & 5.88  & 7.03   & 8.19   & 0.08 & 0.12 & 0.17 & 0.50 \\
  \textcolor{shimmer-main}{$\times$} & L2-hf-70b   & cond. & avg. WTA    & int. & 0.55  & 0.68   & 0.81   & 0.00 & 0.03 & 0.07 & 0.04 \\ \addlinespace[0.5em]
  \textcolor{shimmer-main}{$\times$} & L2-chat-7b  & item  & ---         & prd. & 0.34  & 0.41   & 0.49   & 0.00 & 0.03 & 0.07 & 0.00 \\
  \textcolor{shimmer-main}{$\times$} & L2-chat-7b  & item  & ---         & int. & 0.79  & 1.05   & 1.34   & 0.00 & 0.05 & 0.15 & 0.00 \\
  \textcolor{fern-main}{$\checkmark$} & L2-chat-7b  & cond. & avg. scores & prd. & 58.47 & 69.64  & 81.79  & 0.08 & 0.12 & 0.16 & 0.48 \\
  \textcolor{shimmer-main}{$\times$} & L2-chat-7b  & cond. & avg. scores & int. & 2.03  & 2.37   & 2.73   & 0.00 & 0.02 & 0.05 & 0.00 \\
  \textcolor{shimmer-main}{$\times$} & L2-chat-7b  & cond. & avg. prob.  & prd. & 8.05  & 27.26  & 55.30  & 0.07 & 0.11 & 0.15 & 0.00 \\
  \textcolor{shimmer-main}{$\times$} & L2-chat-7b  & cond. & avg. prob.  & int. & 2.86  & 19.24  & 51.41  & 0.00 & 0.01 & 0.04 & 0.00 \\
  \textcolor{fern-main}{$\checkmark$} & L2-chat-7b  & cond. & avg. WTA    & prd. & 12.18 & 14.70  & 17.21  & 0.08 & 0.12 & 0.16 & 0.50 \\
  \textcolor{shimmer-main}{$\times$} & L2-chat-7b  & cond. & avg. WTA    & int. & 0.86  & 1.01   & 1.14   & 0.00 & 0.02 & 0.05 & 0.00 \\ \addlinespace[0.5em]
  \textcolor{shimmer-main}{$\times$} & L2-chat-13b & item  & ---         & prd. & 0.36  & 0.44   & 0.53   & 0.00 & 0.03 & 0.08 & 0.00 \\
  \textcolor{shimmer-main}{$\times$} & L2-chat-13b & item  & ---         & int. & 0.68  & 0.90   & 1.11   & 0.00 & 0.05 & 0.13 & 0.00 \\
  \textcolor{fern-main}{$\checkmark$} & L2-chat-13b & cond. & avg. scores & prd. & 5.56  & 6.63   & 7.76   & 0.08 & 0.12 & 0.16 & 0.49 \\
